---
title: "EEB498 - Independent Project - Preliminary Data Simulation"
output:
  pdf_document:
    toc: false
    number_section: true
---

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(knitr)
library(FD)
library(TPD)
opts_chunk$set(tidy.opts=list(width.cutoff=70),tidy=TRUE)
ksr_abundance <- read_csv('~/Desktop/Current Courses/EEB498/EEB498_IndpProj/2018_Species_T1_Abundance.csv')
ksr_trait <- read_csv('~/Desktop/Current Courses/EEB498/EEB498_IndpProj/2018_Species_Ontogeny_Trait.csv')
```

**All .csv files are provided by Leila Forsyth (Benjamin Gilbert Lab).**

**Specific leaf area (SLA) will be chosen to be the functional trait of focus (chosen arbitrarily).**

# Preliminary data simulation using Leila's '2018 Ontogeny' dataset

## Manipulation of the dataframes to obtain usable dataframes/matrices

**Had to remove observations for Asclepias_tuberosa, Cirsium_vulgare, Melilotus_albus because there are only one instances of each species.**

```{r}
ksr_abun_true_df <- ksr_abundance %>% 
  select(-c(Plot,Size,Leonurus_cardiaca, Silene_vulgaris, Solanum_dulcamara)) %>% 
  group_by(Block) %>% 
  summarise_each(funs(mean)) %>% 
  column_to_rownames(var='Block')

ksr_abun_true_matrix <- data.matrix(ksr_abun_true_df)

ksr_indiv_trait_df <- ksr_trait %>% 
  select(-c(Plot_Number,Number ,Collection_Number:`Area_(cm2)`)) %>% 
  filter(Species!='Lolium_aparine' & Species!='Silene_latifolia') %>% 
  group_by(Block, Species) %>% 
  mutate(Indiv = seq(1, length(Species), 1))

ksr_mean_trait_true_df <- ksr_indiv_trait_df %>% 
  group_by(Species) %>% 
  summarise(SLA = mean(`SLA_(cm2/g)`)) %>% 
  column_to_rownames(var='Species')

REND_trait_true_df <- ksr_indiv_trait_df %>% 
  filter(Species!='Asclepias_tuberosa' & Species!='Cirsium_vulgare' & Species!='Melilotus_albus')

REND_trait_true_df <- REND_trait_true_df[order(REND_trait_true_df$Species),]

REND_abun_true_df <- ksr_abun_true_df %>% 
  select(-c(Asclepias_tuberosa, Cirsium_vulgare, Melilotus_albus))

REND_abun_true_matrix <- data.matrix(REND_abun_true_df)
```

## Calculation of the "true" value of functional beta diversity indices (CWM, FDis, FRic, FEven) using Leila's '2018 Ontogeny' dataset

https://www.rdocumentation.org/packages/FD/versions/1.0-12/topics/functcomp

https://www.rdocumentation.org/packages/FD/versions/1.0-12/topics/fdisp

https://www.rdocumentation.org/packages/TPD/versions/1.1.0/topics/REND

```{r}
# Community-weighted mean (CWM)

ksr_cwm_true <- functcomp(ksr_mean_trait_true_df, ksr_abun_true_matrix)
(ksr_cwm_true_mean <- mean(ksr_cwm_true$SLA))

# Functional dispersion (FDis)

ksr_fdisp_true <- fdisp(gowdis(ksr_mean_trait_true_df), ksr_abun_true_matrix)
(ksr_fdisp_true_mean <- mean(ksr_fdisp_true$FDis))

# Functional richness and Functional evenness (FRic & FEven)

ksr_TPDs_true <- TPDs(species = REND_trait_true_df$Species, traits = REND_trait_true_df$`SLA_(cm2/g)`)

ksr_TPDc_true <- TPDc(TPDs = ksr_TPDs_true, sampUnit = REND_abun_true_matrix)

(ksr_RicEve_true <- REND(TPDc = ksr_TPDc_true))

(ksr_fric_true_mean <- mean(ksr_RicEve_true$communities$FRichness))
(ksr_feve_true_mean <- mean(ksr_RicEve_true$communities$FEvenness))
```

## Calculation of the "manipulated" value of functional beta diversity indices (CWM, FDis) using Leila's '2018 Ontogeny' dataset

**Errant data treatment for this simulation was removal of 50% of the entire dataset**

Inject the errant data treatment into the trait dataframe.

**Cannot appropriately remove the corresponding abundances within the abundance matrix because the dataframe does not contain the SLA trait values for each individual!!!**

 - not a big issue, because trait data and abundance data are decoupled
 
 - could perform errant data treatments for trait data, abundance data, and for both

```{r}
set.seed(50)

ksr_indiv_trait_df$`SLA_(cm2/g)`[sample(nrow(ksr_indiv_trait_df), round(nrow(ksr_indiv_trait_df)*0.5, digits=0), replace=FALSE)] <- 0

ksr_mean_trait_errant_df <- ksr_indiv_trait_df %>% 
  group_by(Species) %>% 
  summarise(SLA = mean(`SLA_(cm2/g)`)) %>% 
  column_to_rownames(var='Species')

REND_trait_errant_df <- ksr_indiv_trait_df %>% 
  filter(Species!='Asclepias_tuberosa' & Species!='Cirsium_vulgare' & Species!='Melilotus_albus' & Species!='Oenothera_biennis') %>% 
  filter(`SLA_(cm2/g)`!=0)

REND_trait_errant_df <- REND_trait_errant_df[order(REND_trait_errant_df$Species),]

REND_abun_errant_df <- ksr_abun_true_df %>% 
  select(-c(Asclepias_tuberosa, Cirsium_vulgare, Melilotus_albus, Oenothera_biennis, Leucanthemum_vulgare, Verbascum_thapsus))

REND_abun_errant_matrix <- data.matrix(REND_abun_errant_df)

# Community-weighted mean (CWM)

ksr_cwm_manip <- functcomp(ksr_mean_trait_errant_df, ksr_abun_true_matrix[, row.names(ksr_mean_trait_errant_df)])
(ksr_cwm_manip_mean <- mean(ksr_cwm_manip$SLA))

# Functional dispersion (FDis)

ksr_fdisp_manip <- fdisp(gowdis(ksr_mean_trait_errant_df), ksr_abun_true_matrix[, row.names(ksr_mean_trait_errant_df)])
(ksr_fdisp_manip_mean <- mean(ksr_fdisp_manip$FDis))

# Functional richness and Functional evenness (FRic & FEven)

ksr_TPDs_manip <- TPDs(species = REND_trait_errant_df$Species, traits = REND_trait_errant_df$`SLA_(cm2/g)`)

ksr_TPDc_manip <- TPDc(TPDs = ksr_TPDs_manip, sampUnit = REND_abun_errant_matrix)

ksr_RicEve_manip <- REND(TPDc = ksr_TPDc_manip)

(ksr_fric_manip_mean <- mean(ksr_RicEve_manip$communities$FRichness))
(ksr_feve_manip_mean <- mean(ksr_RicEve_manip$communities$FEvenness))
```

## Calculation of the percent relative bias of functional beta diversity indices (CWM, FDis) under the relevant errant data treatment

```{r}
percent_relative_bias <- function(x, y){
  return(((x - y) / y) * 100)
}

# Community-weighted mean (CWM)

(prb_cwm <- percent_relative_bias(x = ksr_cwm_manip_mean, y = ksr_cwm_true_mean))

# Functional dispersion (FDis)

(prb_fdis <- percent_relative_bias(x = ksr_fdisp_manip_mean, y = ksr_fdisp_true_mean))

# Functional richness and Functional evenness (FRic & FEven)

(prb_fric <- percent_relative_bias(x = ksr_fric_manip_mean, y = ksr_fric_true_mean))
(prb_feven <- percent_relative_bias(x = ksr_feve_manip_mean, y = ksr_feve_true_mean))
```

## Depiction of the percent relative bias of functional beta diversity indices (CWM, FDis) under the relevant errant data treatment

```{r}
treatment <- c(0, 1)
perf_cwm <- c(0, prb_cwm)
perf_fdis <- c(0, prb_fdis)
perf_fric <- c(0, prb_fric)
perf_feven <- c(0, prb_feven)
prb_df <- data.frame(treatment, perf_cwm, perf_fdis, perf_fric, perf_feven) %>% 
  pivot_longer(perf_cwm:perf_feven, names_to="index", values_to="prb")

ggplot(prb_df, aes(treatment, prb, colour=index))+
  geom_line(size=1.5)+
  geom_hline(yintercept=0, linetype="dashed")+
  scale_x_continuous(breaks=seq(0, 1, 1), labels=c("0"="Before", "1"="After"))+
  scale_y_continuous(breaks=seq(-100, 100, 5), limits=c(-60, 10))+
  labs(x="Errant data treatment (50% entire dataset)", y="Percent relative bias")
```
